### 2020-05-16 template.tex
Introduction

% This is a general template file for the LaTeX package SVJour3
% for Springer journals. Original by Springer Heidelberg, 2010/09/16
%
% Use it as the basis for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage{authblk}
%
% insert here the call for the packages your document requires
%\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%\usepackage{latexsym}
% etc.
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
\usepackage{latexsym}% 
\usepackage{lineno}% 
\usepackage{amsmath}%
\usepackage{upgreek} % 
\usepackage{mathtools} % 
\usepackage{graphicx}% 
%\usepackage{subcaption}
\usepackage{tikz}% <<<<<<<<<<<< Added by Sadegh
%\linenumbers*[1]% <<<<<<<<<<<< Added by Sadegh
\usepackage{makecell}% <<<<<<<<<<<< Added by Sadegh
\usepackage{natbib}% <<<<<<<<<<<< Added by Sadegh
\usepackage{geometry}% <<<<<<<<<<<< Added by Sadegh
\usepackage{float}
\usepackage{subfigure}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{multicol, blindtext}
\floatstyle{plaintop}
\restylefloat{table}
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Evaluation of UT1 from intensives and 24-hour sessions from CONT17 campaign
%\thanks{}
}
% Grants or other notes about the article that should go on the front
% page should be placed within the \thanks{} command in the title
% (and the %-sign in front of \thanks{} should be deleted)
%
% General acknowledgments should be placed at the end of the article.

% \subtitle{Do you have a subtitle?\\ If so, write it here}

%\titlerunning{Short form of title}        % if too long for running head

\author{Shrishail Raut\textsuperscript{1,2}       \and
        Robert Heinkelmann\textsuperscript{1}     \and
        Sadegh Modiri\textsuperscript{1,2}           \and
        Kyriakos Balidakis\textsuperscript{1}          \and
         Santiago Belda\textsuperscript{3,4}                \and
         Harald Schuh\textsuperscript{1,2}
}

\authorrunning{Shrishail Raut et al}
\institute{Shrishail Subhash Raut \and Robert Heinkelmann \and Sadegh Modiri \and Kyriakos Balidakis \and Harald Schuh \at
             GFZ German Research Centre for Geosciences, Potsdam, Germany \\
             Technische Universit{\"a}t Berlin, Institute for Geodesy and Geoinformation Science, Berlin, Germany
           \and
          Santiago Belda \at
             Image Processing Laboratory (IPL) - Laboratory of Earth Observation (LEO), University of Valencia, Valencia, Spain
             \\
             UAVAC, University of Alicante, Alicante, Spain
              \and
              \textsuperscript{1} GFZ German Research Centre for Geosciences, Potsdam, Germany  \\
              \textsuperscript{2} Technische Universit{\"a}t Berlin, Institute for Geodesy and Geoinformation Science, Berlin, Germany \\ 
              \textsuperscript{3} Image Processing Laboratory (IPL) - Laboratory of Earth Observation (LEO), University of Valencia, Valencia, Spain\\
              \textsuperscript{4}  UAVAC, University of Alicante, Alicante, Spain
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This work deals with validating the established approach of analyzing VLBI intensive sessions to determine dUT1. VLBI sessions from the CONT17 campaign are chosen as they provide continuous VLBI observations over two weeks (28th Nov - 12th Dec 2017) of the currently highest quality. For the standard 24-hour sessions in this campaign, two different legacy networks were involved, the legacy-1 network, which was entirely based on IVS network stations with global distribution, and legacy-2 network involving VLBA and a few IVS network stations for the global extension. In addition to these 24-hour sessions, two different IVS and one Russian intensive sessions were observed every day during CONT17. The dUT1 determined from the intensive sessions are compared with daily and hourly dUT1 from 24-hour sessions during this 15-day time-frame. The results show that the dUT1 determined from intensive sessions do not show good agreement with daily dUT1 from 24-hour sessions; however, it shows better agreement with hourly
dUT1.
\keywords{UT1 \and VLBI \and CONT17}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\section{Introduction}
\label{intro}
The term UT1-UTC represents the corrections to the phase of the Earth rotation angle $\Omega$ UT1 ($\Omega$ is the nominal Earth angular velocity). Very long baseline interferometry is the only space geodetic technique which is can determine UT1-UTC. It is a microwave-based space geodetic technique that measures the difference in signal arrival times from the extragalactic radio source (e.g., the Quasars) receiving simultaneously at two or more radio telescopes. As GNSS satellites require UT1-UTC daily as an important input, this quantity has to be determined fairly on a daily basis. \\
The International VLBI Service for Geodesy and Astrometry (IVS) established so-called Intensive sessions, which are being carried out since 1985. The primary objective of intensive sessions is to determine UT1-UTC, and it consists of one VLBI baseline, and it typically spans over for one hour. The baselines consist of a large east-west extension as UT1 is highly sensitive to it. The stations which take part in the intensive sessions are mentioned in the table. The accuracy of UT1 estimates obtained from intensive sessions is about 15 microseconds, whereas the accuracy of UT1 estimates from 24-hour VLBI sessions is about 6-7 microseconds. The parameterization used for the analysis of intensive sessions are different from the parameterization of 24-hour sessions and can be seen in the table. As seen from the table, we do not estimate most of the parameters while analyzing intensive sessions, because the intensive sessions do not contain enough observations per parameter. Besides, the station coordinates cannot be estimated as one baseline is insufficient to fix the degree of freedom of the terrestrial basis. The dUT1, which is estimated from this approach, may contain inaccuracies. As intensive sessions contain observations for one hour, they give rise to correlations between CPO and terrestrial pole coordinates. Whereas analyzing 24-hour VLBI sessions, the parameters which were fixed to their respective apriori values in the approach mentioned above, are estimated in this approach. This is possible as the 24-hour sessions contain multiple baselines and enough observations per parameter. \\
This can be validated by comparing dUT1 values derived from intensive sessions and 24-hour VLBI sessions. Generally, the 24-hour and intensive sessions coincide only 2-3 days a week, but during continuous VLBI campaign CONT-17, the 24-hour and intensive sessions take place on the same day for 15 days. This provides an excellent opportunity to compare dUT1 values from these sessions. During this campaign, two kinds of 24-hour sessions and three types of the intensive session took place. We will discuss it's detailed in the next section.
\section{Section title}
\label{sec:1}
Text with citations \cite{RefB} and \cite{RefJ}.
\subsection{Subsection title}
\label{sec:2}
as required. Don't forget to give each section
and subsection a unique label (see Sect.~\ref{sec:1}).
\paragraph{Paragraph headings} Use paragraph headings as needed.
\begin{equation}
a^2+b^2=c^2
\end{equation}

% For one-column wide figures use
\begin{figure}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:1}       % Give a unique label
\end{figure}
%
% For two-column wide figures use
\begin{figure*}
% Use the relevant command to insert your figure file.
% For example, with the graphicx package use
  \includegraphics[width=0.75\textwidth]{example.eps}
% figure caption is below the figure
\caption{Please write your figure caption here}
\label{fig:2}       % Give a unique label
\end{figure*}
%
% For tables use
\begin{table}
% table caption is above the table
\caption{Please write your table caption here}
\label{tab:1}       % Give a unique label
% For LaTeX tables use
\begin{tabular}{lll}
\hline\noalign{\smallskip}
first & second & third  \\
\noalign{\smallskip}\hline\noalign{\smallskip}
number & number & number \\
number & number & number \\
\noalign{\smallskip}\hline
\end{tabular}
\end{table}

%\begin{acknowledgements}
%If you'd like to thank anyone, place your comments here
%and remove the percent signs.
%\end{acknowledgements}

% BibTeX users please use one of
%\bibliographystyle{spbasic}      % basic style, author-year citations
%\bibliographystyle{spmpsci}      % mathematics and physical sciences
%\bibliographystyle{spphys}       % APS-like style for physics
%\bibliography{}   % name your BibTeX data base

% Non-BibTeX users please use
\begin{thebibliography}{}
%
% and use \bibitem to create references. Consult the Instructions
% for authors for reference list style.
%
\bibitem{RefJ}
% Format for Journal Reference
Author, Article title, Journal, Volume, page numbers (year)
% Format for books
\bibitem{RefB}
Author, Book title, page numbers. Publisher, place (year)
% etc
\end{thebibliography}

\end{document}
% end of file template.tex